---
title: "\\textbf{HIV Drug Resistance Report}"
output:
  pdf_document:
header-includes:
  - "\\usepackage{fancyhdr}"
  - "\\usepackage{amsfonts,amssymb,amsmath}"
  - "\\usepackage{titlesec}"
  - "\\usepackage{soul}"
  - "\\usepackage{datetime}"
  - "\\usepackage{helvet}"
  - "\\pagestyle{fancy}"
  - "\\fancyhf{}"
  - "\\fancyhead[LE,RO]{\\includegraphics[width=4.0cm]{`r params$logo`}}"
  - "\\fancyfoot[RE,LO]{Report created on \\today}"
  - "\\fancyfoot[LE,RO]{\\thepage}"
  - "\\fancypagestyle{plain}{\\pagestyle{fancy}}"
  - "\\renewcommand{\\familydefault}{\\sfdefault}"
  - "\\renewcommand{\\headrulewidth}{2pt}"
  - "\\renewcommand{\\footrulewidth}{1pt}"
  - "\\titleformat{\\subsection}{\\large\\bfseries\\rlap{\\color{blue!50}\\rule[0.0ex]{\\linewidth}{3ex}\\vspace{-4.2ex}}\\large\\color{white}}{\\thesubsection}{1em}{}"

params:
  dr_report_hivdb: ""
  logo: ""
  dr_comments: ""
  
---

```{r setup, include=FALSE}
library(knitr)
library(plyr)
library(dplyr)
library(jsonlite)
library(tinytex)
library(tidyverse)
#tinytex::install_tinytex()
```

```{r, eval=TRUE, echo=FALSE}
# Import json from sierra-local ----
json_sierra <- jsonlite::fromJSON(params$dr_report_hivdb, flatten = TRUE)

# Create an empty dataframe
df_allDrugResistance <- data.frame()

# Since there is only one sample for this json file
# You may use loop if multiple samples is present in a single json file
index_sample <- 1
df_result <- json_sierra[index_sample,]

sampleName <- df_result$inputSequence.header
df_DR <- df_result$drugResistance[[1]]
geneList <- df_DR$gene.name


# t <- df_result$alignedGeneSequences[[1]]
subtype <- df_result$subtypeText
AApositions <- df_result$alignedGeneSequences[[1]] |>
  select(c(gene.name, firstAA, lastAA))
colnames(AApositions) <- c("Gene", "firstAA", "lastAA")


# Extract Drug Resistance Info ----
for (gene in geneList) {
  
  df_geneDrugResistance <- df_DR |>
    filter(gene.name == gene)
  
  version <- df_geneDrugResistance$version.text
  versionDate <- df_geneDrugResistance$version.publishDate
  
  df_geneDrugScore <- df_geneDrugResistance$drugScores[[1]] |>
    select(c("drugClass.name","drug.name", "score", "text"))
  
  colnames(df_geneDrugScore) <- c("Drug Class", "Drug Name",
                                  "HIVDB Score", "Drug Susceptibility")
  
  df_allDrugResistance <- rbind(df_allDrugResistance, df_geneDrugScore)
}


# Extract Mutations info ----
df_allMutations <- df_result$alignedGeneSequences[[1]] |>
  tidyr::unnest(mutations) |>
  select(c(text, primaryType, gene.name))

colnames(df_allMutations) <- c("Mutation", "Type", "Gene")



# Import the DR comments
df_commentsDR <- read.csv(params$dr_comments) |>
  select(-Type)

# Combine the all mutations of the sample and their corresponding comments
df_allMutationsComments <- left_join(df_allMutations, df_commentsDR,
                                     by = c("Gene", "Mutation")) |>
  drop_na()

```

```{r, echo=FALSE}
# For PR ----
PR_major <- ""
PR_accessory <- ""
PR_other <- ""
PR_firstAA <- ""
PR_lastAA <- ""
PR_comments <- NULL

if ("PR" %in% geneList) {
  PR_major <- df_allMutationsComments |>
    filter(Gene == "PR" & Type == "Major")
  
  PR_accessory <- df_allMutationsComments |>
    filter(Gene == "PR" & Type == "Accessory")
  
  PR_other <- df_allMutationsComments |>
    filter(Gene == "PR" & Type == "Other")
  
  PR_firstAA <- AApositions |>
    filter(Gene == "PR") |>
    pull(firstAA)
  
  PR_lastAA <- AApositions |>
    filter(Gene == "PR") |>
    pull(lastAA)
  
  if (nrow(df_allMutationsComments) == 0 ) {
    PR_comments <- NULL
  } else {
    PR_comments <- df_allMutationsComments |>
      filter(DrugClass == "PI")
  }
}


# For RT ----
RT_NRTI <- ""
RT_NNRTI <- ""
RT_other <- ""
RT_firstAA <- ""
RT_lastAA <- ""
RT_comments <- NULL

if ("RT" %in% geneList) {
  RT_NRTI <- df_allMutationsComments |>
    filter(Gene == "RT" & Type == "NRTI")
  
  RT_NNRTI <- df_allMutationsComments |>
    filter(Gene == "RT" & Type == "NNRTI")
  
  RT_other <- df_allMutationsComments |>
    filter(Gene == "RT" & Type == "Other")
  
  RT_firstAA <- AApositions |>
    filter(Gene == "RT") |>
    pull(firstAA)
  
  RT_lastAA <- AApositions |>
    filter(Gene == "RT") |>
    pull(lastAA)
  
  if (nrow(df_allMutationsComments) == 0 ) {
    RT_comments <- NULL
  } else {
    RT_comments <- df_allMutationsComments |>
      filter(DrugClass == "NRTI" | DrugClass == "NNRTI")
  }
}



# For IN ----
IN_major <- ""
IN_accessory <- ""
IN_other <- ""
IN_firstAA <- ""
IN_lastAA <- ""
IN_comments <- NULL

if ("IN" %in% geneList) {
  IN_major <- df_allMutationsComments |>
    filter(Gene == "IN" & Type == "Major")
  
  IN_accessory <- df_allMutationsComments |>
    filter(Gene == "IN" & Type == "Accessory")
  
  IN_other <- df_allMutationsComments |>
    filter(Gene == "IN" & Type == "Other")
  
  IN_firstAA <- AApositions |>
    filter(Gene == "IN") |>
    pull(firstAA)
  
  IN_lastAA <- AApositions |>
    filter(Gene == "IN") |>
    pull(lastAA)
  
  if (nrow(df_allMutationsComments) == 0 ) {
    IN_comments <- NULL
  } else {
    IN_comments <- df_allMutationsComments |>
      filter(DrugClass == "INSTI")
  }
}

```

``` {r, echo=FALSE}
showList <- function(input) {
  if (nrow(input) > 0) {
    return(input |> pull(Mutation))
  } else {
    return("None")
  }
}

bold_first_word <- function(x) {
  words <- str_split(x, " ", n = 2)
  bolded <- sapply(words, function(y) paste0("**", y[1], "**", ifelse(length(y) > 1, paste0(" ", y[2]), "")))
  return(bolded)
}


showComments <- function(input) {
  if (is.null(input)) {
    return("None")

  } else {
    selected_comment <- input |>
      select(Comment)
    
    knitr::kable(selected_comment, align = "l", escape = FALSE,
                 row.names = FALSE, col.names = NULL)
  }
}

```

## **Sample summary**
**Sample name:** `r sampleName`

**Mapped genes:** `r geneList`

**Subtype:** `r subtype`

**Algorithm version**: `r paste0("HIVDB ", version, " (", versionDate, ")" )`

## **Protease**
**Codon Coverage:** `r paste(PR_firstAA, "-", PR_lastAA)`

**PR Major Mutations:** `r showList(PR_major)`

**PR Accessory Mutations:** `r showList(PR_accessory)`

**Other Mutations:** `r showList(PR_other)`

**Comments:** `r showComments(PR_comments)`

#### **Drug Resistance Score: Protease Inhibitors**
```{r,echo=FALSE}
PR_drugResistance <- df_allDrugResistance |>
  filter(`Drug Class` == "PI") |>
  select(-`Drug Class`)

knitr::kable(PR_drugResistance, align = 'c',
             row.names=FALSE)  
```

\newpage

## **Reverse Transcriptase**
**Codon Coverage:** `r paste(RT_firstAA, "-", RT_lastAA)`

**NRTI-related Mutations**: `r showList(RT_NRTI)`

**NNRTI-related Mutations**: `r showList(RT_NNRTI)`

**Other Mutations**: `r showList(RT_other)`

**Comments**: `r showComments(RT_comments)`

#### **Drug Resistance Score: Nucleoside Reverse Transcriptase Inhibitors**
```{r,echo=FALSE}
RT_NRTI_drugResistance <- df_allDrugResistance |>
  filter(`Drug Class` == "NRTI") |>
  select(-`Drug Class`)

knitr::kable(RT_NRTI_drugResistance, align = 'c',
             row.names=FALSE)  
```

#### **Drug Resistance Score: Non-Nucleoside Reverse Transcriptase Inhibitors**
```{r,echo=FALSE}
RT_NNRTI_drugResistance <- df_allDrugResistance |>
  filter(`Drug Class` == "NNRTI") |>
  select(-`Drug Class`)

knitr::kable(RT_NNRTI_drugResistance, align = 'c',
             row.names=FALSE)  
```

\newpage

## **Integrase**
**Codon Coverage:** `r paste(IN_firstAA, "-", IN_lastAA)`

**IN Major Mutations**: `r showList(IN_major)`

**IN Accessory Mutations**: `r showList(IN_accessory)`

**Other Mutations**: `r showList(IN_other)`

**Comments**: `r showComments(IN_comments)`

#### **Drug Resistance Score: Protease Inhibitors**
```{r,echo=FALSE}
IN_drugResistance <- df_allDrugResistance |>
  filter(`Drug Class` == "INSTI") |>
  select(-`Drug Class`)

knitr::kable(IN_drugResistance, align = 'c',
             row.names=FALSE)  
```
